Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-04-27 overleaf/paper_SIMD.tex
Edits to Analytical Model.

\author{Ana L. Milanova}
\email{milanova@cs.rpi.edu}
\affiliation{%
    \institution{Rensselaer Polytechnic Institute}
    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
    \city{Troy}
    \state{New York}
    %\country{Iceland}
}

@@

In this work we focus on an intermediate language and what we call \emph{circuit-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
in the classical compiler. The following figure summarizes our key idea:

@@

\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific circuit-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code. %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}

@@

Our results emphasize the importance of ``circuit-independent'' optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source. We believe that our work can lead to future work on ``circuit-independent'' compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.

@@

\begin{table*}
\begin{tabular}{ccc}
\begin{minipage}{0.3\textwidth}
{\small
\begin{pythonn}
def biometric(C: shared[list[int]], D: int, 
      S: shared[list[int]], N: int) -> 
      shared[tuple[int,int]]:
   min_sum = MAX_INT
   min_idx = 0  
   for i in range(N): 
      sum = 0
      for j in range(D): 
         # d = S[i,j] - C[j]
         d = S[i * D + j] - C[j] 
         p = d * d
         sum = sum + p
      if sum < min_sum:
         min_sum = sum
         min_idx = i
   return (min_sum, min_idx)
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.33\textwidth}
{\small
\begin{pythonn}
min_sum!1 = MAX_INT
min_idx!1 = 0
# S^ is same as S. C^ replicates C N times: 
S^ = lift(S, ((i * D) + j), (i:N,j:D)) # S^[i,j] = S[i,j]
C^ = lift(C, j, (i:N,j:D)) # C^[i,j] = C[j]

@@

Thus, the fixed cost of the vectorized program amounts to
$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and un-vectorized programs.) 
The first term in the sum corresponds to the vectorized
subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
one corresponds to the remaining FOR loops on $i$. Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
Additionally, the un-vectorized version runs out of memory for N=256, while the vectorized one continues to run up to standard maximal input size N=4096.
Our experimental results illustrate this as well. \ana{Check numbers.}

@@

@@

\begin{enumerate}
\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
instructions have the same cost.
%\ishaq{Don't we need to distinguish between local (cheap) vs. interactive (expensive) instructions? I don't see how we could assume this. -- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
\item We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays. We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
in parallel still cost $2\alpha$. 
%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$. Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays. We have to replace constant costs with functions, as in (3).}
\end{enumerate}

>>>>

\author{Ana Milanova}
\email{milanova@cs.rpi.edu}
\affiliation{%
    \institution{Rensselaer Polytechnic Institute}
    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
    \city{Troy}
    \state{New York}
    %\country{Iceland}
}

@@

In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
in the classical compiler. The following figure summarizes our key idea:

@@

\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code. %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}

@@

Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source. We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.

@@

\begin{table*}
\begin{tabular}{ccc}
\begin{minipage}{0.275\textwidth}
{\small
\begin{pythonn}
def biometric(C: shared[list[int]], D: int, 
      S: shared[list[int]], N: int) -> 
      shared[tuple[int,int]]:
   min_sum = MAX_INT
   min_idx = 0  
   for i in range(N): 
      sum = 0
      for j in range(D): 
         # d = S[i,j] - C[j]
         d = S[i * D + j] - C[j] 
         p = d * d
         sum = sum + p
      if sum < min_sum:
         min_sum = sum
         min_idx = i
   return (min_sum, min_idx)
\end{pythonn}
}
\end{minipage}

@@

\begin{minipage}{0.33\textwidth}
{\small
\begin{pythonn}
min_sum!1 = MAX_INT
min_idx!1 = 0
# S^ is same as S. C^ replicates C N times: 
S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]

@@

Thus, the fixed cost of the vectorized program amounts to
$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and un-vectorized programs.) 
The first term in the sum corresponds to the vectorized
subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
one corresponds to the remaining FOR loops on $i$. Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
\ana{Edit numbers with final experiments.}

@@

\ana{We need an intro to section here.}

@@

\begin{enumerate}
\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
instructions have the same cost.
%\ishaq{Don't we need to distinguish between local (cheap) vs. interactive (expensive) instructions? I don't see how we could assume this. -- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
of $n$ operations costs $\alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution, which costs $n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays. We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
%in parallel still cost $2\alpha$. 
%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$. Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays. We have to replace constant costs with functions, as in (3).}
\end{enumerate}

### 2022-04-27 overleaf/prob_stmt_np_hardness.tex
Edits to Analytical Model.

\ana{Ishaq? Basically, define sequential schedule, then define an equivalent parallel schedule. A parallel schedule is equivalent if it preserves def-use relations in sequential schedule, or in other words, schedules def ahead of the use. Problem is to minimize cost of Parallel schedule.}

At the lowest level, we have two types of MPC instructions (also called \emph{gates} in similar works) 1) local/non-interactive instruction (i.e. an addition instruction $A$) and 2) remote/interactive instruction (i.e. a multiplication instruction $M$). Each instruction in the program is either an $A$-instruction or an $M$-instruction.

@@

Given a serial schedule (a linear graph) of an MPC program i.e. a sequence of instructions $S := (S_1, \dots, S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $S$, our task is to construct a parallel schedule (another linear graph) $P := (P_1, \dots, P_m)$ observing the following conditions:

\begin{enumerate}
    \item Multiple, not necessarily continuous instructions from $S$ can be grouped into a single $P_i$.
    \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e. if instructions $S_i, S_j, i < j$ form a def-use i.e. an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}$ such that $i' < j'$.
\end{enumerate}

\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Preserving def-use \emph{dependencies} means the computed function remains the same in both $S$ and $P$.

In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). We also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g. as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps simplify the problem.}. It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. We use the following cost model:

\begin{enumerate}
    \item $A$ costs $\alpha$ units and $M$ costs $\beta$ unites.
    \item There is unlimited bandwidth i.e. a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions (or $M$-instructions), concretely either $\alpha$ unites or $\beta$ units.
\end{enumerate}

@@

where $cost(S_i) = \alpha$ if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. Similarly, the cost of schedule $P$ is

@@

Each $P_i$ may contain multiple instructions, thus $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions. Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$).

@@

Note that we consider a linearized MPC schedule $S$ above for ease of exposition only. In our tool-chain we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $P$. The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.

@@

As an example, consider the MPC program consisting of the following three sequences, right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node:

\begin{enumerate}
    \item $A \rightarrow M \rightarrow A$
    \item $A \rightarrow A \rightarrow A$
    \item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{\ell'}$ in $P$.

Cost of schedule $P$ is computed using \cref{eq_par_schedule_cost} above.
% The cost of a schedule $P$ is

@@

$$ 
A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3)
$$

The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3). The cost of this schedule is $3\alpha + 2\beta$.

The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978, Vazirani2010}. The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number of total number of nodes.

To see the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node. Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$. It is also a shortest common supersequence. To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$. The cost of $P'$, and $P$, is $X \cdot \alpha + Y \cdot \beta$. Now suppose, there exists a shorter common supersequence, $P''$ that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions. Since $P''$ is shorter than $P'$, therefore $X' + Y' < X + Y$, and $X' \cdot \alpha + Y' \cdot \beta < X \cdot \alpha + Y \cdot \beta$ i.e. $\mathit{cost}(P'') < \mathit{cost}(P')$. But $\mathit{cost}(P') = \mathit{cost}(P)$ and $\mathit{cost}(P)$ is the optimal cost.  Therefore $\mathit{cost}(P'') < \mathit{cost}(P')$ is contradiction and no such $P''$ exists. \qed

\subsection{Loud Thinking}
\ishaq{I will delete this section after we have decided on how to handle various issues raised in this section}.

@@

The problem is that while RHS is the same in \cref{ineq_cost_1} and \cref{ineq_cost_2}, we cannot say anything about the relationship between their LHS. It could be possible that cost of $SS_2$, LHS of  \cref{ineq_cost_2}, is more than cost of $SS_1$, LHS of \cref{ineq_cost_1}. Thus, we have a shorter super sequence $SS_2$ that does not contradict that the optimality of the schedule used to generate $SS_1$.

>>>>

%\ana{Ishaq? Basically, define sequential schedule, then define an equivalent parallel schedule. A parallel schedule is equivalent if it preserves def-use relations in sequential schedule, or in other words, schedules def ahead of the use. Problem is to minimize cost of Parallel schedule.}

As mentioned earlier, at the lowest level, we have two types of MPC instructions (also called \emph{gates} in similar works) 1) local/non-interactive instruction (i.e. an addition instruction $A$) and 2) remote/interactive instruction (i.e. a multiplication instruction $M$). %Each instruction in the program is either an $A$-instruction or an $M$-instruction.

@@

Given a serial schedule (a linear graph) of an MPC program i.e. a sequence of instructions $S := (S_1; \dots; S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $S$, our task is to construct a parallel schedule (another linear graph) $P := (P_1; \dots; P_m)$ observing the following conditions:

\begin{enumerate}
    \item All $P$'s consist of MPC instructions of the same kind, e.g., all MUL, or ADD, etc. 
    \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e. if instructions $S_i, S_j, i < j$ form a def-use i.e. an edge exists from $S_i$ to $S_j$ in $G$, then they can only be mapped to $P_{i'}, P_{j'}$ such that $i' < j'$.
\end{enumerate}

\paragraph{Correctness} Correctness of $P$ is guaranteed by definition. Preserving def-use \emph{dependencies} means the computed function remains the same in both $S$ and $P$.

In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). Recall that we also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g. as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps simplify the problem.}. It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. We use the following cost model:

%\begin{enumerate}
%    \item $A$ costs $\alpha$ units and $M$ costs $\beta$ unites.
%    \item There is unlimited bandwidth i.e. a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions (or $M$-instructions), concretely either $\alpha$ unites or $\beta$ units.
%\end{enumerate}

@@

where $cost(S_i) = \alpha$ or $\beta$. %if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. 
Similarly, the cost of schedule $P$ is

@@

Each $P_i$ may contain multiple instructions, and $\mathit{cost}(P_i)$ is amortized.
Thus, according to our model $\mathit{cost}(P_i) = \alpha_\mathit{fix} + |P_i|\alpha_\mathit{var}$ if $P_i$ stores $\alpha$-instructions, 
or $\mathit{cost}(P_i) = \beta_\mathit{fix} + |P_i|\beta_\mathit{var}$ if it stores $\beta$-instructions.
%if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions. 
Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$), possibly an optimal schedule. 
Originally we hoped that the problem is simper and computation of the optimal schedule is tractable. Unfortunately, the optimal schedule turns out to be NP-hard via a reduction 
to the Shortest Common Supersequence problem.

@@

Note that we consider a linearized MPC schedule $S$ above for ease of exposition only. In our tool-chain we use an MPC Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $P$. %The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.

@@

As an example, consider the MPC program consisting of the following three sequences, all made up of two distinct $\alpha$-instructions $M_1$ and $M_2$. The right arrow indicates a def-use \emph{dependence}, meaning that the source node must execute before the target node:

\begin{enumerate}
    \item $M_1 \rightarrow M_2 \rightarrow M_1$
    \item $M_1 \rightarrow M_1 \rightarrow M_1$
    \item $M_2 \rightarrow M_1 \rightarrow M_2$
\end{enumerate}

%A \emph{schedule} $P: (P_1 ; P_2 \dots ; P_m)$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{\ell'}$ in $P$.

% The cost of a schedule $P$ is

@@

$$ 
M_1(1), M_1(2) \; ;\; M_1(2) \; ; \; M_2(1), M_2(3) \; ; \; M_1(1), M_1(2), M_1(3) \; ; \;  M_2(3)
$$

The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3). 
Cost of schedule $P$ is computed using \cref{eq_par_schedule_cost} above and it amounts to $5\alpha_\mathit{fix} + 9\alpha_\mathit{var}$.

The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978, Vazirani2010}. The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences. For our problem $n$ is the maximum length of a sequence and $k$ is the number of total number of sequences. We can immediately see that the optimal schedule is the shortest schedule, since the shortest schedule minimizes the fixed cost while the variable cost remains the same.

\ana{This paragraph needs work.} To formalize the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
%We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node. 
Clearly, $P'$, which now is a sequence of $M_1$ and $M_2$ nodes, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$. It is also a shortest common supersequence. 
%To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$. 
The cost of $P$, is $L\alpha_\mathit{fix}+N\alpha_\mathit{var}$ where $L$ is the length of $P$ and $N$ is the total number of instructions across all sequences. %$X \cdot \alpha + Y \cdot \beta$.

Now suppose, there exists a shorter common supersequence $P'$. 
that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions. Since $P''$ is shorter than $P'$, therefore $X' + Y' < X + Y$, and $X' \cdot \alpha + Y' \cdot \beta < X \cdot \alpha + Y \cdot \beta$ i.e. $\mathit{cost}(P'') < \mathit{cost}(P')$. But $\mathit{cost}(P') = \mathit{cost}(P)$ and $\mathit{cost}(P)$ is the optimal cost.  Therefore $\mathit{cost}(P'') < \mathit{cost}(P')$ is contradiction and no such $P''$ exists. \qed

\begin{comment}
\subsection{Loud Thinking}
\ishaq{I will delete this section after we have decided on how to handle various issues raised in this section}.

@@

The problem is that while RHS is the same in \cref{ineq_cost_1} and \cref{ineq_cost_2}, we cannot say anything about the relationship between their LHS. It could be possible that cost of $SS_2$, LHS of  \cref{ineq_cost_2}, is more than cost of $SS_1$, LHS of \cref{ineq_cost_1}. Thus, we have a shorter super sequence $SS_2$ that does not contradict that the optimality of the schedule used to generate $SS_1$.
\end{comment}
